Synthesizing Symmetric Lenses

### 2018-06-29 papers/SymmetricLenses/main.tex
notes on new intro

In a big-data world, one will often find similar data stored in a
variety of different formats. For instance, electronic calendars come in iCalendar, vCalendar and h-event
formats~\cite{calendar-formats}; US taxation data can be transmitted via Tax XML (used by the US government
for e-filing) or TXF (used by TurboTax)~\cite{tax-formats}; and sewing machines can read embroidery files
transmitted as PES formats, DST formats or several others~\cite{embroidery-formats}. \dpw{Editorial comment:
Phil Wadler forced his very irritating long-form citations into the standard format. ug. how stupid is that.}

@@

Lens languages, like any domain-specific language, also cut down the space of
possible programs one can write, and the smaller the space of programs,
the easier it is to find the program one is looking for.  
%This feature can be a negative, if the desired transformation
%falls outside of the scope provided by the language, but can also be positive, as complex
%properties (like invertability) can be enforced automatically.
%Another advantage of reducing the
%set of programs is that doing so can facilitate program synthesis, which is essentially
%a search through the space of possible computations that can be expressed as a lens.
Recent research~\cite{optician}
has exploited this fact by demonstrating it is possible to synthesize relatively
large, complex lenses from specifications.
In particular, given types that describe source and target formats, there are synthesis
algorithms that will find bijective lenses~\cite{optician}
for medium-sized files, such as those housing linux
application configurations.  Oftentimes, these algorithms find such lenses in less than
a second, and without the aid of any examples at all, with just types describing the source and
target formats.  \dpw{I would cite the ICFP paper (cite string: maina+:quotient-synthesis) as to appear but I'm not sure how to do that
without possibly violating the spirit of double-blind reviewing. ie, if it hasn't appeared yet,
how could we know the content if we aren't the authors?  Mentioning it in the 3rd person does not
seem to suffice ...}

@@

The first difficulty was pragmatic:  symmetric lenses actually
usually operate over three structures, the source $S$, the target $T$ and a
complement $C$~\cite{symmetric-lenses}.
The latter object contains all the information not present in one of $S$ or $T$.  Logistically,
the complement object must be stored away somewhere and infrastructure must be developed
to manage it.  
Perhaps more relevant though, is that complements complicate specifications, even when
those specifications are just examples.  The issues is that one cannot
just give examples of source and target; one must also give guidance on how to manage
the ``hidden state.''  In order to avoid such complexities, we defined a new
variant of symmetric lenses, which we call \emph{simple symmetric lenses}.  Our new
lenses do not require complements; they only need source and target data as inputs.

@@

With SREs in hand, we develop a new algorithm for synthesizing likely lenses.
This algorithm is constructed from a pair of collaborating search procedures.
Given source and target SREs $S$ and $T$, the first algorithm searches for two \emph{related} SREs
$S'$ and $T'$ that are compatible (a heuristic that estimates the likelihood
our second algorithm will succeed).  $S'$ and $T'$ are related in the sense
that they define the same regular sets as $S$ and $T$ (respectively) and the distribution
of strings is identical.  Indeed, we guarantee these two properties by implementing
a search procedure that obeys the star semi-ring axioms of regular expression equivalence,
and which we have proven does not alter the probability distributions of the languages.

>>>>

In a big-data world, one will often find similar data stored in a
variety of different formats. For instance, electronic calendars come in iCalendar, vCalendar and h-event
formats~\cite{calendar-formats}; US taxation data can be transmitted via Tax XML (used by the US government
for e-filing) or TXF (used by TurboTax)~\cite{tax-formats}; and sewing machines can read embroidery files
transmitted as PES formats, DST formats or several others~\cite{embroidery-formats}. \dpw{Editorial comment:
Phil Wadler forced his very irritating long-form citations into the
standard format. ug. how stupid is that.}
\ksf{I like the embroidery example and I also hate the long form citations.}

@@

Lens languages, like any domain-specific language, also cut down the space of
possible programs one can write, and the smaller the space of programs,
the easier it is to find the program one is looking for.  
%This feature can be a negative, if the desired transformation
%falls outside of the scope provided by the language, but can also be positive, as complex
%properties (like invertability) can be enforced automatically.
%Another advantage of reducing the
%set of programs is that doing so can facilitate program synthesis, which is essentially
%a search through the space of possible computations that can be expressed as a lens.
Recent research~\cite{optician}
has exploited this fact by demonstrating it is possible to synthesize relatively
large, complex lenses from specifications.
In particular, given types that describe source and target formats, there are synthesis
algorithms that will find bijective lenses~\cite{optician}
for medium-sized files, such as those housing linux
application configurations.  \ksf{why medium sized?  You can have very
large files with simple format descriptions that are bijective and for
which our previous systems would have worked just fine.} Oftentimes, these algorithms find such lenses in less than
a second, and without the aid of any examples at all, with just types describing the source and
target formats.  \dpw{I would cite the ICFP paper (cite string: maina+:quotient-synthesis) as to appear but I'm not sure how to do that
without possibly violating the spirit of double-blind reviewing. ie, if it hasn't appeared yet,
how could we know the content if we aren't the authors?  Mentioning it in the 3rd person does not
seem to suffice ...} \ksf{The ICFP 2018 website lists the paper; I
think it is okay to cite it.  The reviewer can assume we contacted the
authors for a copy.  We could ask Stephanie for a ruling if we want to
be sure.  One of the principles of double blind as adopted by SIGPLAN
is not to reduce the quality of the paper; omitting this citation
woudl have that affect.}

@@

The first difficulty was pragmatic:  symmetric lenses actually
usually operate over three structures, the source $S$, the target $T$ and a
complement $C$~\cite{symmetric-lenses}.
The latter object contains all the information not present in one of $S$ or $T$.  Logistically,
the complement object must be stored away somewhere and infrastructure must be developed
to manage it.  
Perhaps more relevant though, is that complements complicate specifications, even when
those specifications are just examples.  The issues is that one cannot
just give examples of source and target; one must also give guidance on how to manage
the ``hidden state.''  In order to avoid such complexities, we defined a new
variant of symmetric lenses, which we call \emph{simple symmetric lenses}.  Our new
lenses do not require complements; they only need source and target
data as inputs. \ksf{This seems magic.  Say something about how it is
possible/under what circumstances it is possible to not need a
complement?
At the price of some expressiveness, our new lenses eliminate the need
for complements; they require only source and target data as inputs.
}

@@

With SREs in hand, we develop a new algorithm for synthesizing likely lenses.
This algorithm is constructed from a pair of collaborating search procedures.
Given source and target SREs $S$ and $T$, the first algorithm searches for two \emph{related} SREs
$S'$ and $T'$ that are \emph{compatible} (a heuristic that estimates the likelihood
our second algorithm will succeed).  $S'$ and $T'$ are related in the sense
that they define the same regular sets as $S$ and $T$ (respectively) and the distribution
of strings is identical.  Indeed, we guarantee these two properties by implementing
a search procedure that obeys the star semi-ring axioms of regular expression equivalence,
and which we have proven does not alter the probability distributions of the languages.
